---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Lecture Notes
When curve dips down rapidly there is new technology

##Hands-on Session
###OMIM
Search "asthma Verlaan" and click first result and scroll down until you find "Verlaan et. al" then click the SNPs and click where it says it overlaps with `4 transcripts` 
Ensemble oage gives info on SNP and the genes in its vicinity (genomic context) and tells you populations in various regions 

###Galaxy
FASTQC runs all quality scores in box plots 
Tophat reads over the introns in data for RNA-seq

###Analyze RNA-seq by genotype results

Here we read the results of RNA-Seq for about ~230 samples (people) with different genotypes for the rs8067378 Asthma associated SNP.

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
head(expr)
#Calculate average expression value of people with A/A by extracting the A/A rows
inds <- expr$geno=="A/A"
expr[inds,]
mean(expr[inds,"exp"])
summary(expr[inds,]$exp)
```

```{r}
#Do same thing for G/G
indg <- expr$geno=="G/G"
summary(expr[indg,]$exp)
```

```{r}
boxplot(exp ~ geno, data = expr, notch=TRUE)
#if notches don't overlapp then there's statistic significance 
```

