---
title: 'Class 6: R Functions'
author: "Sarra Larif"
date: "1/23/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Functions

## Level 2 Heading

### Level 3 Heading

```{r}
#this is a silly plot
plot(1:5)
```
Any text in R markdown is just text, need to do insert (cntrl+alt+i) to add R code

Lets see more about file import (i.e. reading files into R). The main read function in base R is `read.table()`

`read.csv` has different defaults 

```{r}
t1 <- read.csv("test1.txt")
t1
```

```{r}
t2 <- read.table("test2.txt", sep = "$", header = TRUE)
t2
```

```{r}
t3 <- read.table("test3.txt")
t3
```

```{r}
add <- function(x, y = 1) {
  #sum the input x and y
  x + y
}
add(5)
```
Green play button sends code to "R brain"

Let's use the new function
```{r}
add(175, 3)
```

How does this work with vector inputs?
```{r}
add(c(5, 7, 2, 9))
```
It just added 1 to each.

```{r}
add(c(1, 2, 4), 4)
```
It added 4 to each. 

```{r}
add(c(1, 2, 4), c(1, 2, 4))
```

```{r}
#add(c(1,2,4), 1, 2, 4) brings an error:unusued arguments
```

What is this `range()` function?
```{r}
x <- c(4,4,10,3,11)
max(x)
min(x)
range(x)
```

```{r}
rescale <- function(x) {
 rng <-range(x)
 (x - rng[1]) / (rng[2] - rng[1])
}
rescale(x)
```

Do some more testing
```{r}
rescale(1:10)
```

How to get function to work here where there is an NA
```{r}
rescale(c(1, 2, NA, 3, 10))
```
```{r}
x <- c(1, 2, NA, 3, 10)
rng <-  range(x, na.rm = TRUE)
rng
```
```{r}
rescale2 <- function(x) {
 rng <-range(x, na.rm = TRUE)
 (x - rng[1]) / (rng[2] - rng[1])
}
rescale2(x)
```

```{r}
rescale3 <- function(x, na.rm=TRUE, plot=FALSE) {
 rng <-range(x, na.rm=na.rm)
 print("Hello")
 answer <- (x - rng[1]) / (rng[2] - rng[1])
 print("is it me you are looking for?")
 if(plot) {
 plot(answer, typ="b", lwd=4)
 }
 print("I can see it in ...")
 return(answer)
}

rescale3(x)
```

```{r}
rescale3 <- function(x, na.rm=TRUE, plot=FALSE) {
 rng <-range(x, na.rm=na.rm)
 print("Hello")
 answer <- (x - rng[1]) / (rng[2] - rng[1])
 print("is it me you are looking for?")
 if(plot) {
   print("Don't sing again please!")
 plot(answer, typ="b", lwd=4)
 }
 print("I can see it in ...")
 return(answer)
}

rescale3(x, plot = TRUE)
```

Do not put install commands in markdown or it will try to install the package every time you render the file

Can you improve this code?
```{r}
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```
Plot is measure of flexibility along sequence from c to n (higher peak = more it's moving)

Questions: 
Q1. `read.pdb()` provides a file in the global environment that gives data from PDB on the protein
It's a large list of 8 things and of class "pdb", "sse"
```{r}
class(s1)

str(s1)
```

```{r}
s1$seqres
```
A/B = chain A sequences and chain B sequences 

```{r}
aa321(s1$seqres)
```

Q2. `trim.pdb()` provides the same data as read.pdb but with less information (subset of data)
"produce a new smaller PDB object, containing a subset of atoms, from a given larger PDB object"

Q3. removing `sse` removes the grey rectangles in the plot which represents what the secondary structure of the protein is (helix, sheet, turn)

Q4. Combining the plots
```{r}
plotb3(s1.b, sse=s1.chainA, typ="l", lwd = 2, ylab="Bfactor")
points(s2.b, typ = "l", col = "blue", lwd = 2)
points(s3.b, typ = "l", col = "green", lwd = 2)
```

Q5. 2 and 3
```{r}
hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
plot(hc)
```
Quantify with `dist( rbind(s1.b, s2.b, s3.b) )` which provides the distances between each protein