---
title: "Class 6 Homework"
author: "Sarra Larif"
date: "1/23/2020"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Improve this code: 
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")

```{r}
#Input the PDB ID with optional specifications for the chain and atom of interest as well as for the plot type and label
plot.prot <- function(PDB_ID, chain = "A", atom = "CA", typ = "l", ylab = "Bfactor"){
 
   #library attaches the necessary package to allow access of PDB data
  library(bio3d) 
  
  #read.pdb uses the PDB ID input to call the PDB file of that protein
  a <- read.pdb(PDB_ID)  
  
  #trim.pdb narrows the information of the PDB file to only provide what is needed for the plot while including for any specifications made in the new function for the chain or atom
  b <- trim.pdb(a, chain = chain, elety = atom)
  
  #c stores the selection of data for the plot the function calls
  c <- b$atom$b
 
   #plotb3 calls the output which is a scatter plot of the selected protein structure residues including secondary structure element representations in the margins
  plotb3(c, sse = b, typ = typ, ylab = ylab)
}
```

Examples of output: 
```{r}
plot.prot("4AKE")
plot.prot("1AKE")
plot.prot("1E4Y")
```

