---
title: "Class 12: Structural Bioinformatics Part 1"
author: "Sarra Larif"
date: "2/13/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Notes: 
Sequence: hydrophobic amino acids want to be buried/hidden while hydrophilic want to be on surface so amino acid chain is dynamic (shape-shifting molecule)with different shapes/conformations active at different times 
Low energy shapes/conformations are stable
Structure is more conserved than sequence because it's closer to function so even when sequence changes, the shape needs to stay the same to keep the function 
###Protein Structure
1IEP is a kinase that is a drug target for leukemia 
ABDGE next to the element is for carbon (alpha, beta, etc)
Tube representation just draws through every alpha carbon of each protein
NMA gives us better idea of how proteins work to help us better devlop drugs for them

##PDB Statistics

Here we inspect the types of structures in the main database for 3D biomolecular data - the PDB

>Q1: Determine the percentage of structures solved by X-Ray and Electron Microscopy. 

```{r}
#read in file
stats <- read.csv("Data Export Summary.csv", row.names = 1)
stats

#calculate percentage values asked for each 'method
ans <- stats$Total/sum(stats$Total)*100
names(ans) <- rownames(stats)
round(ans,2)
ans

sum(stats$Proteins)

perc.xray <- stats[1,1]/stats[1,5]
perc.EM <- stats[3,1]/stats[3,5]
perc.xray
perc.EM

sum(stats$Total)

perc.xe <- (stats[1,1]+stats[3,1])/sum(stats$Total)*100
perc.xe
```

>Q2. Also can you determine what proportion of structures are protein? Aim to have a rendered GitHub document with working code that yields your answers. 

```{r}
round(sum(stats$Proteins)/sum(stats$Total)*100,2)
```

>Q3: Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?

1289

##VMD
Need to download PDB format of the molecule then change "determine file type" to automatically to open 
to change appearance: graphics -> representations, change it then -> apply -> create rep
Resolution of PDB structure is 2 angstroms but hydrogen is smaller than that so they don't appear in water molecules that "hover" around the protein 
To take high res image: File->Render->tachyon internal-> render

```{r}
library(bio3d) 
pdb <- read.pdb("1hsg")
pdb
```

##select protein only atoms

Here I will use the `atom.select()` function 
```{r}
prot <- atom.select(pdb, "protein", value = TRUE)
lig <- atom.select(pdb, "ligand", value = TRUE)
```

and write out these new PDB objects with `write.pbd()`
```{r}
write.pdb(lig, file = "1hsg_ligand.pdb")
write.pdb(prot, file = "hsg1_protein.pdb")
```

