---
title: "Class 07"
author: "Sarra Larif"
date: "1/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source("http://tinyurl.com/rescale-R")
```

```{r}
function(x, na.rm=TRUE, plot=FALSE, ...) {
  # Our rescale function from lecture 6
  if( !is.numeric(x) ) {
    stop("Input x should be numeric", call.=FALSE)
  }
  rng <-range(x, na.rm=TRUE)
  answer <- (x - rng[1]) / (rng[2] - rng[1])
  if(plot) { 
    plot(answer, ...) 
  }

  return(answer)
}
```

#Warning/Stop Functions
`warning` gives a warning text and keeps the function going
`stop` gives a warning message and breaks the code 
`!` in front of variable means not i.e. `!TRUE = FALSE`
In the rescale function above: rescale only works on numbers so `!is.numeric` gives if the input is NOT numeric (i.e. purple) then it will return TRUE and the `stop` but if it is numeric, it will give FALSE and skip the `stop`
```{r}
is.numeric(40)
!is.numeric(40)
```

```{r}
x <-  c(1, 2, NA, 3, NA)
y <-  c(NA, 3, NA, 3, 4)
typeof(x)
is.numeric(x)
is.na(x)
typeof(is.numeric(x))
z <- c("dvnsivf", 2, NA, 3, NA)
is.na(z)
#is.na works for characters too
```


```{r}

is.both.na <- function (x,y) {
is.na(x) & is.na(y)
}
is.both.na(x, y)

a <- c(4, 5, NA, 8)
b <- c(NA, 9, NA, 18, 2)
is.both.na(a, b)

is.both.na1 <-  function(x,y) {
  if(length(x) != length(y)){
    stop("your inputs aren't the same length")
  }
  is.na(x) & is.na (y)
}
is.both.na1(a, b)
```

```{r}
rescale_new <-  function(x, y) {
  ## Print some info on where NA's are as well as the number of them 
  if(length(x) != length(y)) {
    stop("Input x and y should be vectors of the same length", call.=FALSE)
  }
  na.in.both <- ( is.na(x) & is.na(y) )
  na.number  <- sum(na.in.both) #Gives how many NAs are there
  na.which   <- which(na.in.both) #Gives position of NAs

  message("Found ", na.number, " NA's at position(s):", 
          paste(na.which, collapse=", ") ) 
  
  return( list(number=na.number, which=na.which) )
}
rescale_new(x,y)
```

```{r}
truth_vector <- c(FALSE, TRUE, FALSE, TRUE)
which(truth_vector)
```

#Data Frames
Data frames are tables that allow you to do special functions on 
```{r}
df1
df1[1,1]
df1[1,]
```
First variable is row, second is column

Can look at row/column headers 
```{r}
df1[, "IDs"]
df1$IDs
```

Can clean up global space
```{r}
rm(x)
rm(y)
rm(z)
rm(a)
rm(b)
rm(truth_vector)
```

Look at difference between IDs in 2 data frames
```{r}
x <- df1$IDs
y <- df2$IDs

x
y 

intersect(x,y) #which genes are in BOTH data sets?

x %in% y #which elements of x exist in y?
x[x %in% y] #subsets variables in x that are also in y

y %in% x
y[y %in% x]
```
 We want to make a matrix of x elements in y and y elements in x
```{r}
  cbind(x[ x %in% y], y[y %in% x])
```
 Can make a function out of something we already typed by highlighting that code and then pressing code on top bar and pressing extract function or control+alt+x
```{r}
gene_interect <- function(x, y) {
  cbind(x[ x %in% y], y[y %in% x])
}

gene_intersect(x,y)
```
 
 Want to take data frames and have the function intersect for us
```{r}
#Need to change x and y to the data frames so it works with any data frame not just the ones we stored into the variables 
gene_interect2 <- function(df1, df2) {
  cbind(df1[df1$IDs %in% df2$IDs, ], df2[df2$IDs %in% df1$IDs, ], "exp") 
  #Add comme after %in% function within bracket to get all information associated with what is within the data
}

x <- df1$IDs
y <- df2$IDs

```
 
Want to add functionality 
```{r}
gene_interect3 <- function(df1, df2, gene.colname = "IDs") {
  
  cbind(df1[df1[, gene.colname] %in% df2[, gene.colname],], df2[df2[, gene.colname] %in% df1[, gene.colname], ], "exp")
  
}

#These are all the same and we want to add this functionality so we can subset with different columns in future
df1$IDs
df1[, "new_name"]
gene.colname = "new_name"
df1[, gene.colname]

colnames(df1) <- c("new_name", "exp")
colnames(df2) <- c("new_name", "exp")
#change column name in data frame

gene_intersect3(df1, df2, gene.colname = "new_name")
```
 
 Cleaning up the code
```{r}
gene_interesect4 <- function(df1, df2, gene.colname = "IDs") {
  
  df1.name <- df1[, gene.colname]
  df2.name <- df2[, gene.colname]
  
  df1.inds <- df1.name %in% df2.name
  df2.inds <- df2.name %in% df1.name
  
  cbind(df1[df1.inds,], df2[df2.inds,], "exp")
}

gene_intersect4(df1, df2, gene.colname = "new_name")
```
 
#Installing Packages
`install.packages('ggplot2')`
```{r}
library("ggplot2")
```
R graph gallery on google shows you all the different graphs you can make in R
`ggplot` only takes data frames 
```{r}
dim(iris) #a built-in data set; dim = dimensions -> 150 x 5 dataframe
head(iris)
levels(iris$species)

#Specifying how to present data visually
myplot <- ggplot(data = iris, aes(x=sepal.length, #Data on x axis
                                   y=sepal.width)) #Data on y axis
class(myplot)

ggplot(data = iris, aes(Sepal.Length, Sepal.Width, color = Species)) + geom_point((aes(shape = Species)), size = 3) +geom_smooth(method = "lm")
#make sure to add + because ggplot adds data layer by layer
#color/shape = species plots the data of species by color/shape
# lm = linear modeling; shaded region is confidence interval 
```

Bar Plot
```{r}
ggplot(iris, aes(Sepal.Length, fill = Species)) + geom_bar(stat = "count")
#only need one variable
#alpha = transparency 
# can also add vertical lines for medians 
#can change background
ggplot(iris)
```

rstudio.com/cheatsheet gives cheatsheet for different ggplots
 
#Group Work
