---
title: "Class 13: Structure based drug discovery"
author: "Sarra Larif"
date: "2/18/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Applications
Discriminate between good and poor binders, or provide a priority ranking to a collection of ligands
Provide in depth mechanistic characterization of specific ligant or group of ligands
Provide valuable guidance for medicinal chemists trying 
###computational ligand docking
Screening and ranking compounds as potential ligands (aka virtual screening)
Improving "lead" compounds (aka ligand optimization, more on this later...)
  This is a common practice among seasoned computational chemists
###Scenario I
Want a drug that fits better than the peptide to block it 
Use docking software and look for LOWER energy 
Want to find best fit docking based on energy interactions
Bonding Interactions vs. non-binding interactions 

##Hands-On Section

Download and clean HIV-Pr Receptor Structure

Here we will download the 1HSG PDB structure from the online database, remove the bound drug from Merk and write out a protein only and ligand only new set of PDB files

We will use the **bio3d** package functions `get.pdb()`, `read.pdb()`, `atom.select()`, and `write.pdb()`

`library(bio3d)`
```{r}
#Download PDB file
file.name <- get.pdb("1hsg")

pdb <- read.pdb(file.name)
pdb
```

Select Protein only
```{r}
prot <- atom.select(pdb, string= "protein" , value = TRUE)
write.pdb(prot, file = "1hsg_protein.pdb")
```

Select and write out ligand only file
```{r}
lig <- atom.select(pdb, string = "ligand", value = TRUE)
write.pdb(lig, file = "1hsg_ligand.pdb")
lig
```

###AutoDockTools
Add hydrogens with edit>hydrogens>add
Grid>macromolecules>choose>1hsg_protein>okay>save in class 
pdbqt is input for function
Grid>Grid Box: only want ligands to dock at a certain binding site it limits the center of mass so they can only go there (???)
Change size of center of mass to make sure ligand can fit within that box (more for visualization to show where ligand is contrained to)

Section 1-1.5
Took pdb file and split it to protein and ligand then added hydrogens and parameters (atom types and charges - .pdbqt)

Make text file with config info for the autodock software to read

```{r}
res <- read.pdb("all.pdbqt", multi=TRUE)
write.pdb(res, "results.pdb")
```

Best fit result is the first position displayed on VMD

###Ligand Based
Uses ligands that have already been discovered/established
