---
title: "Class 10"
author: "Sarra Larif"
date: "2/6/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Scaling
Sometimes we use different scales to compare things (meters vs. ft)
`prcomp(x, SCALE = TRUE)` scales the data before PCA (normally scale = false so it doesn't scale)

##Analysis of Human Breast Cancer
```{r}
#Import the data from the link in the class website and store as wisc.df
wisc.df <- read.csv("WisconsinCancer.csv")
head(wisc.df)
```
M = malignant, B = benign 
Column X is full of NAs (problem)
ID column is zip codes that aren't supposed to be there because good neighborhoods showed treatment success and bad neighborhoods didn't
Don't want Diagnosis either because that's the answer and don't want X
```{r}
wisc.data <- as.matrix(wisc.df[,3:32])
head(wisc.data)
```
```{r}
#how many patients in this data set?
nrow(wisc.data)
#how many Ms and Bs?
table(wisc.df$diagnosis)
#how many variables/features in data are suffixed with _mean?
grep("_mean", colnames(wisc.df), value=TRUE)
#value = TRUE tells you what they are 
length(grep("_mean", colnames(wisc.df)))
```

##Principal Component Analysis 
Before we use PCA we need to consider if we need to scale `scale = TRUE`
```{r}
colMeans(wisc.data)
apply(wisc.data,2,sd) #apply standard deviation across (1 for rows or 2 for columns) in data set
```
Want to rounf everything to 2 sig figs
```{r}
round(apply(wisc.data,2,sd),2)
```
Need to scale
```{r}
wisc.pr <- prcomp(wisc.data, scale = TRUE)
summary(wisc.pr)
```
Q4. proportion of variance = can plot 44% of data in PC1
 - The 1st PC captures 44.27% of the original variance.
looking for point where there's  the biggest drop in percent variance exlpained (scree plot?)
Q5. 3 PCS
Q6. 7 PCs

Let's make some figures (visualize PCA results)
```{r}
biplot(wisc.pr) 
#wtf is this? we need to clean it up
```

```{r}
#Tells you the different results we have in PCA
attributes(wisc.pr)
```
```{r}
#plot x of PCA 1 vs x of PCA 2 coloring by if its malignant or benign
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=wisc.df$diagnosis)
```
Red and black are split but red is more spread out and there is a small mixture of malignant and benign but there is a spread so we should look into it more 
(0,0) is mean of all the columns
Tight clustering of black means there aren't many ways to look benign (black) but there are many ways to look malignant (red)
Maximum variation of data is between the minimum and maximum of data (line along dots on page 21 of slides)
take midpoint as zero and genes on end can have plus and minus values so that absolute value of higher number means more variation/higher influence
PC1 and PC2 are each an axis(?)
Each point represents a column of data by plotting its influence on PC1 vs. its influence on PC2
getting points near each other (clustering) means they behave similarly 
```{r}
#Add average lines to visualize easier 
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=wisc.df$diagnosis)
abline(h=0, col="gray", lty=2)
abline(v=0, col="gray", lty=2)
```
malignant and benign are fairly similar and cluster around average?


###Cluster in PC space 
can you take the raw data and cluster it? Yes but it looks terrible 
Let's see if we can do it
```{r}
km <- kmeans(wisc.data, centers = 2, nstart = 20)
hc <- hclust(dist(wisc.data))
plot(hc)
```
Uneven distribution because the data doesn't really have good means of clustering 
instead of raw data, let's give it the results of the principal component analysis
Using the minimum number of principal components required to describe at least 90% of the variability in the data, create a hierarchical clustering model with the linkage method="ward.D2". We use Ward’s criterion here because it is based on multidimensional variance like principal components analysis. Assign the results to wisc.pr.hclust.
```{r}
wisc.pr.hclust <- hclust(dist(wisc.pr$x[,1:3]), method = "ward.D2")
plot(wisc.pr.hclust)
```
To get out clusters out of this tree we need to CUT it with `cutree()`
```{r}
grps3 <- cutree(wisc.pr.hclust, k=3)
grps3
table(grps3) #how many data points are in each group
```

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=grps3)
```
green seems to be benign and black and red seem to be malignant
Let's check: we can use the `table()` function to compare $diagnosis vector with out cluster results vector 
```{r}
table(grps3, wisc.df$diagnosis)
```

```{r}
grps2 <- cutree(wisc.pr.hclust, k=2)
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=grps2)
table(grps2, wisc.df$diagnosis)
```
in cluster 1 you have 24 benign patients but 333 benign patients in cluster 2; malignant has 179 in cluster 1 and 33 in cluster 2(24 and 33 misdiagnosed)
Balance of sensitivity vs. specificity 

##Prediction
We will use the `predict()` function that will take our PCA model from before and new cancer cell data and project that into our PCA space
```{r}
new <- read.csv("new_samples.csv")
new
```
Use `predict()` with our previous PCA model and new data...
```{r}
npc <- predict(wisc.pr, newdata = new)
npc
```
Now draw the PCA plot and add new data:
```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=wisc.df$diagnosis)
points(npc[,1], npc[,2], col ="blue", pch=15, cex=3)
text(npc[,1], npc[,2], labels = c(1,2), col="white")
```
Want to prioritize patient 2 for checks because matches malignant clustering